\pkg{Rgbp}: Bayesian Hierarchical Modeling and Frequentist Method Check

### 2013-04-07 jss/article.tex
writing journal theory almost done

\subsection[Normal-Normal]{Normal-Normal}
\code{gbp} can construct a two-level Normal-Normal hierarchical model on the 8 school data. For reference,  $\sigma^{2}_{j}$ below is assumed to be known or to be accurately estimated, and subscript \emph{j} indicates \emph{j}-th school in the dataset.
\begin{displaymath}
y_{j}\vert \mu_{j} \stackrel{ind}{\sim}\textrm{Normal}(\mu_{j}, \sigma^{2}_{j}),
\end{displaymath}
\begin{displaymath}
\mu_{j}\vert \beta, A\stackrel{ind}{\sim}\textrm{Normal}(\mu_{0j}, A),
\end{displaymath}

where $\mu_{0j} =x_{j}'\beta,~j=1, \ldots, 8$, and a paramter with a zero sub-index represents it is a mean parameter of the prior (second-level) distribution. Based on this conjugate prior distribution, it is easy to derive corresponding posterior distribution. 
\begin{displaymath}
\mu_{j}\vert \textbf{y}, \beta, A \stackrel{ind}{\sim}\textrm{Normal}(~(1-B_{j})y_{j} + B_{j}\mu_{0j},~(1-B_{j})\sigma^{2}_{j}),
\end{displaymath}
where $B_{j}\equiv\frac{\sigma^{2}_{j}}{\sigma^{2}_{j} + A},~j=1, \ldots, 8$ is called a shrinkages and we will explain it in the following section.

\subsection[Poisson-Gamma]{Poisson-Gamma}
\code{gbp} is also able to build a Poisson-Gamma multi-level model on the 31 hospital data. Note that a constant multiplied to the notation representing Gamma distribution is a scale and a square bracket below indicates [mean, variance] of distribution.
\begin{displaymath}
z_{j}\vert \lambda_{j} \stackrel{ind}{\sim}\textrm{Poisson}(n_{j}\lambda_{j}),
\end{displaymath}
\begin{displaymath}
\lambda_{j}\vert \beta, r\stackrel{ind}{\sim}\frac{1}{r}\textrm{Gamma}(\lambda_{0j}r)\sim \textrm{Gamma}[\lambda_{0j}, ~\frac{\lambda_{0j}}{r}],
\end{displaymath}

where $\log(\lambda_{0j}) =x_{j}'\beta$, and $j=1, \ldots, 31$. Immediate posterior distribution of this Poisson-Gamma model is
\begin{displaymath}
\lambda_{j}\vert \textbf{z}, \beta, r \stackrel{ind}{\sim}\frac{1}{r + n_{j}}\textrm{Gamma}(r\lambda_{0j} + z_{j})\sim\textrm{Gamma}[\lambda^{\ast}_{j},~\frac{\lambda^{\ast}_{j}}{r+n_{j}}],
\end{displaymath}

@@

\subsection[Binomial-Beta]{Binomial-Beta}
Binomial-Beta hierarchical model is the last model that \code{gbp} can fit. Again, a square bracket below indicates [mean, variance] of distribution.
\begin{displaymath}
z_{j} \vert p_{0j}\stackrel{ind}{\sim}\textrm{Binomial}(n_{j}, ~p_{j}),
\end{displaymath}
\begin{displaymath}
p_{j} \vert \beta, r\stackrel{ind}{\sim}\textrm{Beta}(rp_{0j},~ r(1-p_{0j}))\sim \textrm{Beta}[p_{0j}, ~\frac{p_{0j}(1-p_{0j})}{r + 1}],
\end{displaymath}

where $\log(\frac{p_{0j}}{1-p_{0j}}) =x_{j}'\beta~\textrm{and}~j=1, \ldots, 18$. Then posterior distribution is
\begin{displaymath}
p_{j}\vert \textbf{z}, \beta, r \stackrel{ind}{\sim}\textrm{Beta}(rp_{0j}+z_{j},~r(1-p_{0j})+n_{j}-z_{j})\sim\textrm{Beta}\bigg[p^{\ast}_{j},~ \frac{p^{\ast}_{j}(1-p^{\ast}_{j})}{r+n_{j}+1}\bigg],
\end{displaymath}
where $p^{\ast}_{j}\equiv(1-B_{j})(\frac{z_{j}}{n_{j}})+B_{j}p_{0j},~ \textrm{and}~ B_{j}\equiv\frac{r}{r+n_{j}}$, $j=1,\ldots,18$.

\subsection[Hyper-prior Distribution]{Hyper-prior Distribution}
Hyper-prior distribution indicates a distribution of the second-level parameters, which plays an important role in deriving a full posterior distribution of all the parameters. \code{gbp} sets a non-informative distribution on second-level parameters to let the data speak more about their estimation.
\begin{displaymath}
\beta \sim \textrm{flat on}~ \mathbf{R}^{m},~~A ~(\textrm{or}~ \frac{1}{r})\sim \textrm{Unif}(0, \infty),
\end{displaymath}

@@

Once we are sure that the hierarchical modeling can add a huge value on our research, the next natural quetsion will be about the validity of interval estimates. ``Does the confidence interval obtained via this Bayesian model-fitting process achieve pre-specified 100($1-\alpha$)\% confidence level for any true parameter values?'' Our answer is ``yes'' and \pkg{Rgbp} has a function to assure this point.

\subsection{Pseudo-data Generation Process}
\subsection{Unbiased Coverage Probability}
\subsection{Rao-Blackwellized Coverage Probability}
\section[Example]{Example}
\subsection[Known Prior Mean]{Known Prior Mean}
Using 10 hospital data with PR
\subsection[Unknown Prior Mean and No Covariate]{Unknown Prior Mean and No Covariate}
Using 8 school data with GR
\subsection[Unknown Prior Mean and One Covariate]{Unknown Prior Mean and One Covariate}
Using 18 baseball data with BR

>>>>

\subsection[Normal-Normal]{Normal-Normal}
\code{gbp} can construct a two-level Normal-Normal hierarchical model on the 8 school data. For reference,  $\sigma^{2}_{j}$ below is assumed to be known or to be accurately estimated, and subscript \emph{j} indicates \emph{j}-th school in the dataset.
\begin{equation}
y_{j}\vert \mu_{j} \stackrel{ind}{\sim}\textrm{Normal}(\mu_{j}, \sigma^{2}_{j}),
\end{equation}
\begin{equation}
\mu_{j}\vert \beta, A\stackrel{ind}{\sim}\textrm{Normal}(\mu_{0j}, A),
\end{equation}

where $\mu_{0j} =x_{j}'\beta,~j=1, \ldots, 8$, and a paramter with a zero sub-index represents it is a mean parameter of the prior (second-level) distribution. Based on this conjugate prior distribution, it is easy to derive corresponding posterior distribution. 
\begin{equation}
\mu_{j}\vert \textbf{y}, \beta, A \stackrel{ind}{\sim}\textrm{Normal}(~(1-B_{j})y_{j} + B_{j}\mu_{0j},~(1-B_{j})\sigma^{2}_{j}),
\end{equation}
where $B_{j}\equiv\frac{\sigma^{2}_{j}}{\sigma^{2}_{j} + A},~j=1, \ldots, 8$ is called a shrinkages and we will explain it in the following section.

\subsection[Poisson-Gamma]{Poisson-Gamma}
\code{gbp} is also able to build a Poisson-Gamma multi-level model on the 31 hospital data. Note that a constant multiplied to the notation representing Gamma distribution is a scale and a square bracket below indicates [mean, variance] of distribution.
\begin{equation}
z_{j}\vert \lambda_{j} \stackrel{ind}{\sim}\textrm{Poisson}(n_{j}\lambda_{j}),
\end{equation}
\begin{equation}
\lambda_{j}\vert \beta, r\stackrel{ind}{\sim}\frac{1}{r}\textrm{Gamma}(\lambda_{0j}r)\sim \textrm{Gamma}[\lambda_{0j}, ~\frac{\lambda_{0j}}{r}],
\end{equation}

where $\log(\lambda_{0j}) =x_{j}'\beta$, and $j=1, \ldots, 31$. Immediate posterior distribution of this Poisson-Gamma model is
\begin{equation}
\lambda_{j}\vert \textbf{z}, \beta, r \stackrel{ind}{\sim}\frac{1}{r + n_{j}}\textrm{Gamma}(r\lambda_{0j} + z_{j})\sim\textrm{Gamma}[\lambda^{\ast}_{j},~\frac{\lambda^{\ast}_{j}}{r+n_{j}}],
\end{equation}

@@

\subsection[Binomial-Beta]{Binomial-Beta}
Binomial-Beta hierarchical model is the last model that \code{gbp} can fit. Again, a square bracket below indicates [mean, variance] of distribution.
\begin{equation}
z_{j} \vert p_{0j}\stackrel{ind}{\sim}\textrm{Binomial}(n_{j}, ~p_{j}),
\end{equation}
\begin{equation}
p_{j} \vert \beta, r\stackrel{ind}{\sim}\textrm{Beta}(rp_{0j},~ r(1-p_{0j}))\sim \textrm{Beta}[p_{0j}, ~\frac{p_{0j}(1-p_{0j})}{r + 1}],
\end{equation}

where $\log(\frac{p_{0j}}{1-p_{0j}}) =x_{j}'\beta~\textrm{and}~j=1, \ldots, 18$. Then posterior distribution is
\begin{equation}
p_{j}\vert \textbf{z}, \beta, r \stackrel{ind}{\sim}\textrm{Beta}(rp_{0j}+z_{j},~r(1-p_{0j})+n_{j}-z_{j})\sim\textrm{Beta}\bigg[p^{\ast}_{j},~ \frac{p^{\ast}_{j}(1-p^{\ast}_{j})}{r+n_{j}+1}\bigg],
\end{equation}
where $p^{\ast}_{j}\equiv(1-B_{j})(\frac{z_{j}}{n_{j}})+B_{j}p_{0j},~ \textrm{and}~ B_{j}\equiv\frac{r}{r+n_{j}}$, $j=1,\ldots,18$.

\subsection[Hyper-prior Distribution]{Hyper-prior Distribution}
Hyper-prior distribution indicates a distribution of the second-level parameters, which plays an important role in deriving a full posterior distribution of all the parameters. \code{gbp} sets a non-informative distribution on second-level parameters to let the data speak more about their estimation.
\begin{equation}
\beta \sim \textrm{flat on}~ \mathbf{R}^{m},~~A ~(\textrm{or}~ \frac{1}{r})\sim \textrm{Unif}(0, \infty),
\end{equation}

@@

Once we are sure that the hierarchical modeling can add a huge value on our research, the next natural quetsion will be about the validity of interval estimates. ``Does the confidence interval obtained via this Bayesian model-fitting process achieve pre-specified 100($1-\alpha$)\% confidence level for any true parameter values?'' Our answer is ``yes'' and \pkg{Rgbp} has a function to assure this point. From now on, all the explanations will be based on the Binomial model.

\subsection{Pseudo-data Generation Process}
Figure 1 will be helpful to understand this process. As we can see in (8), the distribution of each true parameter ($p_{j},~j=1,\ldots, 18$) depends on two hyper-parameters, $r$ and $\beta_{(m\times1)}$, where $m$ is the number of regression coefficients. Once we fix these hyper-parameters at specific values, we can generate true parameters. Suppose we sampled 500 $\textbf{p}_{(18\times1)}$'s, \emph{i.e.}, $\{\textbf{p}^{(i)}_{(18\times1)},~i=1, \ldots, 500\}$. Then, we can also generate $\{\textbf{z}^{(i)}_{(18\times1)},~i=1, \ldots, 500\}$ given each $p_{ij},$ where $i$ indicates $i$-th pseudo-dataset and $j$ does $j$-th player. Next, \code{coverage} fits the Binomial hierarchical model 500 times to obtain 500*18 interval estimates. Taking the first player as an example, we have 500 pairs of ($z_{i1}$, $p_{i1}$), $j=1,\ldots, 500$ and 500 interval estimates.
\begin{figure}[h]
\begin{center}
\includegraphics[width=5cm]{process.png}
\caption{Pseudo-data generating process}
\end{center}
\end{figure}

\subsection{Unbiased Coverage Probability}
Based on the generated pseudo-datasets, let's define an indicator variable, $I_{ij}$, which is 1 if $j$-th interval estimate from $i$-th pseudo-dataset includes $p_{ij}$ and 0 otherwise. One way to estimate the coverage probability is to take average over those indicator variables. We call it an unbiased coverage probability estimate. For example, $\bar{I}_{1}=\sum_{i=1}^{500}I_{i1}/500$ is the estimated coverage probability for the first player.
\subsection{Rao-Blackwellized Coverage Probability}
Based on the definition of Rao-Blackwellization, we define $E(I_{ij}\vert z_{ij}, r, \beta$) as the Rao-Blackwellized coverage probability, where $r$ and $\beta$ were fixed at first (see 5.1). This expectation is the same as P$(\hat{p}_{ij, low}\le p_{ij} \le\hat{p}_{ij, upp}\vert z_{ij}, r, \beta)$, where $\hat{p}_{ij, low}$ and $\hat{p}_{ij, upp}$ are interval estimates (numeric values). We can calculated this probability exactly because we know the distribution of $(p_{ij} \vert z_{ij}, r, \beta)$ is the Beta distribution in (9). For instance, the first player's Rao-Blackwellized coverage probability can be estimated by $\sum_{i=1}^{500}E(I_{i1}\vert z_{i1}, r, \beta)/500$.
\section[Example]{Example}
\subsection[Known Prior Mean]{Known Prior Mean}
Using 10 hospital data with PR
\subsection[Unknown Prior Mean and No Covariate]{Unknown Prior Mean and No Covariate}
Using 8 school data with GR
\subsection[Unknown Prior Mean and One Covariate]{Unknown Prior Mean and One Covariate}
Using 18 baseball data with BR
